---
title: "Covid 19 data"
output: html_document
date: "2025-04-26"
---
## Importing packages
```{r packages}
library(tidyverse)
library(lubridate)
```

## Inputting files

### Inputting links
```{r links}
url_in="https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/"
file1="time_series_covid19_confirmed_global.csv"
file2="time_series_covid19_deaths_global.csv"
file3="time_series_covid19_confirmed_US.csv"
file4="time_series_covid19_deaths_US.csv"

file_names <- c(file1, file2, file3, file4)
urls <- str_c(url_in, file_names)

uid_lookup_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/UID_ISO_FIPS_LookUp_Table.csv"
uid <- read_csv(uid_lookup_url) %>% select(-c(Lat, Long_, Combined_Key, code3, iso2, iso3, Admin2))
```

### Inputting files
```{r import_data, message= FALSE}
global_cases <- read_csv(urls[1])
global_deaths <- read_csv(urls[2])
US_cases <- read_csv(urls[3])
US_deaths <- read_csv(urls[4])
global_cases_ind <- read_csv(urls[1])
global_deaths_ind <- read_csv(urls[2])

```

## Processing

### processing global_cases
```{r processing_global_cases}
global_cases <- global_cases %>% 
  pivot_longer(cols = -c(`Province/State`,
                         `Country/Region`,Lat, Long), names_to = "date", values_to = "cases") %>% select(-c(Lat,Long))
```

### processing global_deaths
```{r processing_global_deaths}
global_deaths <- global_deaths %>% pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), names_to = "date", values_to = "deaths") %>% select(-c(Lat, Long))
```

### joining both data
```{r joining_both}
global <- global_cases %>% full_join(global_deaths) %>% rename(Country_Region = `Country/Region`, Province_State = `Province/State`) %>% mutate(date = mdy(date))

global <- global %>% filter(cases > 0)

global <- global %>% unite("Combined_Key", c(Province_State, Country_Region), sep = ",", na.rm =TRUE, remove = FALSE)
```

### processing US_cases
```{r US}
US_cases <- US_cases %>% pivot_longer(cols = -(UID:Combined_Key), names_to = "date", values_to = "cases") %>% select(Admin2:cases) %>% select(Admin2:cases) %>% mutate(date = mdy(date)) %>% select(-c(Lat, Long_))

US_deaths <- US_deaths %>% pivot_longer(cols = -(UID:Population), names_to = "date", values_to = "deaths") %>% select(Admin2:deaths) %>% mutate(date = mdy(date)) %>% select(-c(Lat, Long_))

US <- US_cases %>% full_join(US_deaths)

```

### Joining with population
```{r join_pop}
global <- global %>% left_join(uid, by = c("Province_State", "Country_Region")) %>% select(-c(UID,FIPS)) %>% select(Province_State, Country_Region, date, cases, deaths, Population, Combined_Key)
```

### Analysis
```{r analysis}

US_by_state <- US %>% group_by(Province_State, Country_Region, date) %>% summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>% mutate(deaths_per_mill = deaths*1000000/Population) %>% select(Province_State, Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% ungroup()

US_totals <- US_by_state %>% group_by(Country_Region, date) %>% summarize(cases = sum(cases), deaths = sum(deaths), Population = sum(Population)) %>% mutate(deaths_per_mill = deaths * 1000000/Population) %>% select(Country_Region, date, cases, deaths, deaths_per_mill, Population) %>% ungroup()

```

## Plots

### global
```{r pressure}
US_totals %>% filter(cases > 0) %>% ggplot(aes(x = date, y = cases)) + geom_line(aes(color = "cases")) + geom_point(aes(color="cases")) + geom_line(aes(y = deaths, color = "deaths")) + geom_point(aes(y = deaths, color = "deaths")) + scale_y_log10() +theme(legend.position="bottom", axis.text.x = element_text(angle = 90)) + labs(title = "COVID19 in US", y= NULL)
```

### New York
```{r plot_new_york}
state <- "New York"

US_by_state %>% filter(Province_State == state) %>% filter(!is.na(cases), cases > 0, deaths > 0) %>% ggplot(aes(x = date, y = cases)) + geom_line(aes(color = "cases")) + geom_point(aes(color="cases")) + geom_line(aes(y = deaths, color = "deaths")) + geom_point(aes(y = deaths, color = "deaths")) + scale_y_log10() +theme(legend.position="bottom", axis.text.x = element_text(angle = 90)) + labs(title = "COVID19 in US", y= NULL)
```

## Deep dive
```{r deep_dive}
US_by_state <- US_by_state %>% mutate(new_cases = cases - lag(cases), new_deaths = deaths - lag(deaths))

US_totals <- US_totals %>% mutate(new_cases = cases - lag(cases), new_deaths = deaths - lag(deaths))

US_totals %>% filter(new_cases > 0, new_deaths > 0) %>% ggplot(aes(x = date, y = new_cases)) + geom_line(aes(color = "new_cases")) + geom_point(aes(color="new_cases")) + geom_line(aes(y = new_deaths, color = "new_deaths")) + geom_point(aes(y = new_deaths, color = "new_deaths")) + scale_y_log10() +theme(legend.position="bottom", axis.text.x = element_text(angle = 90)) + labs(title = "COVID19 in US", y= NULL)

```

### New York
```{r new_york_deep_dive}
US_by_state %>% filter(Province_State == state) %>% filter(new_cases > 0, new_deaths > 0) %>%ggplot(aes(x = date, y = new_cases)) + geom_line(aes(color = "new_cases")) + geom_point(aes(color="new_cases")) + geom_line(aes(y = new_deaths, color = "new_deaths")) + geom_point(aes(y = new_deaths, color = "new_deaths")) + scale_y_log10() +theme(legend.position="bottom", axis.text.x = element_text(angle = 90)) + labs(title = "COVID19 in US", y= NULL)
```

### Worst and best cases
```{r worst_and_best}
US_state_totals <- US_by_state %>% group_by(Province_State) %>% summarize(deaths = max(deaths), cases = max(cases), population = max(Population), cases_per_thou = 1000* cases/population, deaths_per_thou = 1000*deaths/population) %>% filter(cases > 0, population >0)

US_state_totals %>% slice_min(deaths_per_thou, n=10)

US_state_totals %>% slice_min(deaths_per_thou, n =10) %>% select(deaths_per_thou, cases_per_thou, everything())

US_state_totals %>% slice_max(deaths_per_thou, n =10) %>% select(deaths_per_thou, cases_per_thou, everything())

```

### Modelling
```{r models}
# Fit the model
mod <- lm(deaths_per_thou ~ cases_per_thou, data = US_state_totals)

# Create a grid of x values (spread over the range of your data)
x_grid <- seq(min(US_state_totals$cases_per_thou), max(US_state_totals$cases_per_thou), length.out = 300)

# Predict on the new grid
new_df <- tibble(cases_per_thou = x_grid) %>%
  mutate(pred = predict(mod, newdata = .))

# Plot
ggplot(US_state_totals, aes(x = cases_per_thou, y = deaths_per_thou)) +
  geom_point(color = "blue") +
  geom_line(data = new_df, aes(x = cases_per_thou, y = pred), color = "red") +
  labs(title = "Deaths vs. Cases per Thousand",
       x = "Cases per 1,000 people",
       y = "Deaths per 1,000 people")

```

### Looking into India

```{r India_data}
indcases <- global_cases_ind %>% filter(`Country/Region` == 'India')
inddeaths <- global_deaths_ind %>% filter(`Country/Region` == 'India')
ind_cases <- indcases %>% pivot_longer(cols = -c(`Province/State`,                       `Country/Region`,Lat, Long), names_to = "date", values_to = "cases") %>% select(-c(Lat,Long))

ind_deaths <- inddeaths %>% pivot_longer(cols = -c(`Province/State`, `Country/Region`, Lat, Long), names_to = "date", values_to = "deaths") %>% select(-c(Lat, Long))

ind <- ind_cases %>% full_join(ind_deaths)
ind <- ind %>% select(-'Province/State')

#Converting to month
ind$date <- as.Date(ind$date, format = "%m/%d/%Y")
ind$dmonth <- month(ind$date)
ind$dyear <- year(ind$date)

ind$smonth <- as.numeric(ind$dmonth)
ind$syear <- as.character(ind$dyear)
ind$smonth <- sprintf("%02d", ind$smonth)

ind$ym1 <- paste(ind$syear, ind$smonth, sep = "-")

ind <- ind %>%
  mutate(ym2 = ym(ym1))

#Summarising
ind_my <- ind %>%
  filter(!is.na(dyear), !is.na(dmonth)) %>%
  group_by(ym) %>%
  summarise(
    cases = sum(cases, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(cases > 0, deaths > 0)
```

## Plots

### plot of India - cumulative

```{r indplot_cumulative}
ind_my %>%  ggplot(aes(x = ym, y = cases)) + geom_point(aes(color="cases")) + geom_point(aes(y = deaths, color = "deaths")) + scale_y_log10() +theme(legend.position="bottom", axis.text.x = element_text(angle = 90)) + labs(title = "COVID19 in India", y= NULL)
```

### plot of India - new cases
``` {r indplot_newcases}
ind_my <- ind_my %>% mutate(new_cases = cases - lag(cases), new_deaths = deaths - lag(deaths))

ind_my %>% filter(new_cases > 0, new_deaths > 0) %>% ggplot(aes(x = ym, y = new_cases)) + geom_point(aes(color="new_cases")) + geom_point(aes(y = new_deaths, color = "new_deaths")) + scale_y_log10() +theme(legend.position="bottom", axis.text.x = element_text(angle = 90)) + labs(title = "COVID19 in US", y= NULL)
```

### Predictive model
``` {r model_India}

ind_my <- ind_my %>% filter(new_cases > 0, new_deaths > 0)

mod <- lm(new_deaths ~ new_cases, data = ind_my)

x_grid <- seq(min(ind_my$new_cases), max(ind_my$new_cases), length.out = 300)

new_df <- tibble(new_cases = x_grid) %>%
  mutate(pred = predict(mod, newdata = .))

# Plot
ggplot(ind_my, aes(x = new_cases, y = new_deaths)) +
  geom_point(color = "blue") +
  geom_line(data = new_df, aes(x = new_cases, y = pred), color = "red") +
  labs(title = "New Cases vs New Deaths with Regression Line",
       x = "New Cases", y = "New Deaths")

```